---
title: "Notebook, KDHEtools"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Package creation helper code chunks.

# Package

Package related code.

## Create Package
Use the code below to create the package.  Assumes you are in the development package directory

```{r, eval=FALSE}
# NEWS
# Render then Copy NEWS so picked up in help
rmarkdown::render("NEWS.rmd", "all")
file.copy("NEWS.md", "NEWS", overwrite = TRUE)
file.remove("NEWS.html")
#file.remove("NEWS.md")
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Library Name
myPkg <- "KDHEtools"
# Load Library
library(devtools)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create Package
# create(myLibrary)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# files to library folder
# file.copy("metric.values.R","./Library/MMIcalc/R/metric.values.R",overwrite=TRUE)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Document, Install, and Reload Library
## Generate Documentation
#setwd(file.path(".", myPkg))
devtools::document()
## Install New Package (locally)
setwd("..") # return to root directory first
devtools::install(myPkg, quick=FALSE, reload=TRUE, build_vignettes = TRUE)

# blank row above allows for the code to continue instead of waiting for user response to pkg updates.
## Reload library
library(myPkg, character.only = TRUE)
# change wd back to package
setwd(paste0("./", myPkg))
# Help page
help(package=(myPkg))
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

After creating the package reload it after restarting R within RStudio (Control+Shift+F10).

```{r, eval=FALSE}
# Restart R within RStudio:  Ctrl + Shift + F10
myPkg <- "KDHEtools"
library(myPkg, character.only = TRUE)
help(package=(myPkg))
#?metric.scores 
citation(myPkg)
```

## PkgDown website
Don't always need a github website but when needed use Pkgdown.
```{r PkgDown, eval=FALSE}
# Create PkgDown Website
usethis::use_pkgdown()
# Update the files
pkgdown::build_site()
```


## Vignette
Can build separately or when build the entire package.

devtools::build() more useful and build_vignettes()

```{r, eval=FALSE}
# generate Vignette
library(KDHEtools)
library(devtools)
devtools::build_vignettes()
```

## Including Data
All data should have already been created.

use \data-raw\ to store raw data files and the scripts to create the data files 
for internal use in the package.

## Error Checking
Need to use testthat but not to that stage yet.
```{r, eval=FALSE}
# Check for errors (or press Cmd + Shift + E in RStudio)
#http://r-pkgs.had.co.nz/check.html
devtools::check()
```

# Testings

https://testthat.r-lib.org/

https://r-pkgs.org/tests.html

Set up folders with `devtools::test()`.  This is also the command to run tests.

Create test file `usethis::use_test("name")`
