---
title: "Instructions and Background Information"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
always_allow_html: true
output:
  html_fragment:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results='asis', echo=FALSE, warning=FALSE, message = FALSE)
```

# Index Calculator Instructions

WDEQtools was developed to calculate diatom metrics and Indices of Biotic Integrity (IBI) for perennial, wadeable streams in Wyoming. Detailed documentation of the Wyoming Diatom IBI development can be downloaded [here](https://github.com/Blocktt/ShinyAppDocuments/tree/main/WDEQtools/Reports). Users can run any data through the Index calculator and get a result. However, if samples do not meet the criteria listed on the Background page, results should be interpreted with caution.

The Instructions are divided into three sections: 1) preparing the input file, 2) Index Calculator operation, and 3) frequently asked questions (FAQ). Links to an example input file and a document containing more detailed information on preparing the input file are also provided.

## Preparing the input file

WDEQtools requires input files to be in a tabular format, saved as either comma-separated values (CSV) (most commonly used) or tab-separated (.txt or .tsv) files. Table 1 contains a list of required fields. Column names must be capitalized and match the spelling and symbology exactly as shown (some column headings have underscores). Note, diatom traits are not required for each record (i.e., trait fields allow null values). Users can obtain the standard diatom nomenclature as well as all required traits (a combination of the Wyoming Diatom Master Taxa Table and USGS traits database) from Attachment 1 of the Wyoming Diatom IBI development report found [here](https://github.com/Blocktt/ShinyAppDocuments/tree/main/WDEQtools/Reports).

Click here to view an [EXAMPLE INPUT FILE](https://github.com/Blocktt/ShinyAppDocuments/tree/main/WDEQtools/Example_Files). If the user is unsure how to download a data file from GitHub, [watch this short video](https://www.youtube.com/watch?v=GIJdfuAoqFI) or contact Eric Hargett ([eric.hargett\@wyo.gov](mailto:eric.hargett@wyo.gov){.email}) or Jeremy Zumberge ([jeremy.zumberge\@wyo.gov](mailto:jeremy.zumberge@wyo.gov){.email}).

#### Table 1. List of required fields. Column names must be capitalized and match the spelling and symbology exactly as shown.

```{r table_1}
library(readxl)
library(knitr)
library(kableExtra)
# state directories
table.dir <- "tables"
table.file <- "Instruction_Tables.xlsx"
tab1.dir <- "Instr_Table_Diatoms"

table1 <- read_excel(file.path(table.dir, table.file), sheet = tab1.dir
                     , na = c("NA", ""), trim_ws = TRUE, skip = 0
                     , col_names = TRUE)  

# kable(table1)
table1 %>% 
  kbl() %>% 
  kable_styling(full_width = F, position = "left", font_size = 14)
```

### Metric Adjustment Factors

The Wyoming Diatom IBI contains four metrics that are adjusted based on site-specific StreamCat estimates [(StreamCat)](https://www.epa.gov/national-aquatic-resource-surveys/streamcat-dataset). To retrieve StreamCat data for a site, the user must first obtain the appropriate COMID by conducting a spatial join between the site location (i.e., latitude/longitude) and the NHDPlus V2 flowline layer (outside of the app). Once each input site has been assigned a COMID, the user can use the **"Metric Adjustment Download"** tab to obtain COMID-specific StreamCat data. Metric adjustment factor data are required for the Index Calculator input file. See Table 2 for additional information on the StreamCat variables. See Table 3 for necessary input fields. For more information on the derivation of these metric adjustment factors, please read the [Wyoming Diatom IBI report](https://github.com/Blocktt/ShinyAppDocuments/tree/main/WDEQtools/Reports). Note, only COMIDs from the state of Wyoming can be used. If the user would like COMIDs from other, nearby states, please contact Eric Hargett ([eric.hargett\@wyo.gov](mailto:eric.hargett@wyo.gov){.email}) or Jeremy Zumberge ([jeremy.zumberge\@wyo.gov](mailto:jeremy.zumberge@wyo.gov){.email}) and they can update the available list. **Site elevation (ft)** is also a metric adjustment factor, but must be provided by the user.

#### Table 2. Background information on StreamCat variables

```{r table_2}
library(readxl)
library(knitr)
library(kableExtra)
# state directories
table.dir <- "tables"
table.file <- "Instruction_Tables.xlsx"
tab2.dir <- "StreamCatVars"

table2 <- read_excel(file.path(table.dir, table.file), sheet = tab2.dir
                     , na = c("NA", ""), trim_ws = TRUE, skip = 0
                     , col_names = TRUE)  

# kable(table2)
table2 %>% 
  kbl() %>% 
  kable_styling(full_width = F, position = "left", font_size = 14)
```

#### Table 3. List of required fields. Column names must be capitalized and match the spelling and symbology exactly as shown.

```{r table_3}
library(readxl)
library(knitr)
library(kableExtra)
# state directories
table.dir <- "tables"
table.file <- "Instruction_Tables.xlsx"
tab3.dir <- "MetAdjFactInput"

table3 <- read_excel(file.path(table.dir, table.file), sheet = tab3.dir
                     , na = c("NA", ""), trim_ws = TRUE, skip = 0
                     , col_names = TRUE)  

# kable(table3)
table3 %>% 
  kbl() %>% 
  kable_styling(full_width = F, position = "left", font_size = 14)
```

#### Metric Adjustment Download Instructions

Once opened, the user will see the metric adjustment download interface. The user should follow the onscreen instructions as follows:

1.  Load file
    -   **Default settings work well**. If the user is unsure of how to format the data, simply store as a .csv and use the default load settings.
    -   An example input file can be downloaded from Github [EXAMPLE LINK](https://github.com/Blocktt/ShinyAppDocuments/tree/main/WDEQtools/Example_Files)
    -   Choose Separator. The separator indicates how the data is stored. Comma indicates the input file is a comma-separated file (.csv). Tab indicates the input file is a tab-separated file (.txt or .tsv). Finally, Semicolon indicates the input file is a semicolon-separated file which is uncommon in the United States but common in Western-European countries. Be certain that the designated separator is not used in any of the text fields.
    -   Choose Quote. The quote indicates how the data is stored. Double quote is the most common.
    -   Choose file to upload. Hit the browse button and search for the input file.
    -   Once uploaded, make sure data is correct in the viewer. If the incorrect Separator or Quote is chosen, the user may receive an error or the data may look incorrect in the viewer.
2.  Obtain factors
    -   Click 'Obtain factors' button.
    -   A progress bar will appear in the bottom, right corner of the screen. Results cannot be downloaded (step 3) until the progress bar is completed and disappears.
3.  Download results
    -   Select the button to download a zip file with inputs and results.
    -   Check 'results_log_maf.txt' for any warnings or messages. Note, some warnings are automatically generated by R. Feel free to reach out to Eric Hargett ([eric.hargett\@wyo.gov](mailto:eric.hargett@wyo.gov){.email}) or Jeremy Zumberge ([jeremy.zumberge\@wyo.gov](mailto:jeremy.zumberge@wyo.gov){.email}) for any questions related to warnings.
4.  Combine metric adjustment factor data with taxonomic data for reupload into **Index Calculator**.
    -   The downloaded metric adjustment factor data will need to be merged with taxonomic data (outside of the app) to properly calculate the IBI. For reference, see the structure of the example files [HERE](https://github.com/Blocktt/ShinyAppDocuments/tree/main/WDEQtools/Example_Files).

## Index Calculator Operation Instructions

Once the 'Index Calculator' tab is opened, the user will see the Index Calculator interface. The user should follow the onscreen instructions as follows:

1.  Load file
    -   **Default settings work well**. If the user is unsure of how to format the data, simply store as a .csv and use the default load settings.
    -   An example input file can be downloaded from Github [EXAMPLE LINK](https://github.com/Blocktt/ShinyAppDocuments/tree/main/WDEQtools/Example_Files)
    -   Choose Separator. The separator indicates how the data is stored. Comma indicates the input file is a comma-separated file (.csv). Tab indicates the input file is a tab-separated file (.txt or .tsv). Finally, Semicolon indicates the input file is a semicolon-separated file which is uncommon in the United States but common in western European countries. Be certain that the designated separator is not used in any of the text fields (e.g., for multiple BC_USGS values, separate with a semicolon).
    -   Choose Quote. The quote indicates how the data is stored. Double quote is the most common.
    -   Choose file to upload. Hit the browse button and search for the input file.
    -   Once uploaded, make sure data look correct in the viewer. If the incorrect Separator or Quote is chosen, the user may receive an error or the data may look incorrect in the viewer.
2.  Calculate Index
    -   Click the 'Calculate Metric Values and Scores' button.
    -   A progress bar will appear in the bottom, right corner of the screen. Results cannot be downloaded (step 3) until the progress bar is completed and disappears.
3.  Download results
    -   Select the button to download a zip file with inputs and results.
    -   Check 'results_log.txt' for any warnings or messages. Note, some warnings are automatically generated by R. Feel free to reach out to Eric Hargett ([eric.hargett\@wyo.gov](mailto:eric.hargett@wyo.gov){.email}) or Jeremy Zumberge ([jeremy.zumberge\@wyo.gov](mailto:jeremy.zumberge@wyo.gov){.email}) for any questions related to warnings.

## Frequently asked questions and troubleshooting

1.  Why am I getting an error saying that I am missing columns even when I am not?
    -   You may have incorrectly spelled a given column. Try writing the column in all capital letters. Also, some columns (e.g., INDEX_REGION) require an underscore to separate the two words.
2.  Why does my data look strange in the data viewer?
    -   You likely have the incorrect Separator or Quote selected. Otherwise, there may be commas in text fields when the comma separator is selected.
3.  The Index calculation is taking forever to calculate, has the app frozen?
    -   Even though R works remarkably fast, a large dataset will cause the app to slow down. The estimate is approximately 30 seconds for small datasets; however, this could extend to a few minutes for large datasets.
