---
title: "Background Input Metrics"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
always_allow_html: true
output:
  html_fragment:
    toc: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results='asis', echo=FALSE, warning=FALSE, message = FALSE)
```

## Input metrics and scoring formulae

There are five input metrics in the KS MMI (Table 1). Each has its own
random forest model and set of predictor variables. Predictor variables
are used in the random forest analysis to account for natural gradient
effects on the metrics. For each site, the random forest model looks at
site characteristics through decision \'trees\' and arrives at a
predicted metric values for that site based on the site characteristics,
as captured by the predictor values.

The random forest adjusted metric value = **observed value -- predicted
value**

The residual (adjusted) value can be thought of as the variation
remaining after accounting for natural factors.

The adjusted metric residuals were standardized to a 100-point scale
such that higher values represented better conditions. The metric
standardization was an interpolation of metric values between the 5th
and 95th percentiles of each metric distribution, using the scoring
formulas in Table 1. If metric values met the reference expectation
(predicted = observed), they received scores in the 50-60 range. MMI
values were calculated as the average of the individual metric scores.

#### Table 1. Metrics and scoring formulae for the Kansas Macroinvertebrate MMI. All metrics were adjusted by randomForest models (RF adj) and then scored based on the adjusted values. 5th/95th are the percentiles of metric values that were used for metric scoring.

```{r table_1}
library(readxl)
library(knitr)
library(kableExtra)
# state directories
table.dir <- "tables"
table.file <- "Instruction_Tables.xlsx"
tab1.dir <- "BugMetrics"

table1 <- read_excel(file.path(table.dir, table.file), sheet = tab1.dir
                     , na = c("NA", ""), trim_ws = TRUE, skip = 0
                     , col_names = TRUE)  

options(knitr.kable.NA = '')
table1 %>%
  kable() %>%
  kable_styling(full_width = F, position = "left", font_size = 14)
```
