---
title: "Assign Index Class, About"
date: "2023-07-28"
output: html_fragment
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE
                      , results = 'asis'
                      , warning = FALSE
                      , message = FALSE)
```

# Purpose

Each KS MMI input metric has its own random forest model and set of
predictor variables that account for natural gradient effects. There are
12-16 predictor variables per metric (Table 1). Most came from the USEPA
StreamCat dataset (Hill et al. 2016), which includes measures of
geology, soils, hydrology and climate. Other predictor variables include
Omernik Level 3 ecoregion, longitude and modeled median flow (CFS; Perry
et al. 2004). Some of the predictor variables are difficult for some
users to access. We added this function to make that step easier. It
utilizes R tools from
[StreamCat](https://github.com/USEPA/StreamCatTools) and
[NHDPlus](https://doi-usgs.github.io/nhdplusTools/) to add in the
predictor variables based on the latitude and longitude coordinates in
the input file.

```{r table_1}
library(readxl)
library(knitr)
library(kableExtra)
# state directories
table.dir <- "tables"
table.file <- "Instruction_Tables.xlsx"
tab1.dir <- "MetAdjFactors"

table1 <- read_excel(file.path(table.dir, table.file), sheet = tab1.dir
                     , na = c("NA", ""), trim_ws = TRUE, skip = 0
                     , col_names = TRUE)  

# kable(table2)
options(knitr.kable.NA = '')

table1 %>% 
  kbl() %>% 
  kable_styling(full_width = F, position = "left", font_size = 14)
```

![Assign predictors
workflow](images%5CFileBuilder_Predictors_Fig1.png){width="80%"}

# Input file

The following fields are required:

-   SampleID or SiteID -- unique identifier

    -   Tip: pick an identifier that can be linked to other files (such
        as the file coming out of the File Builder -- Taxa
        Translator/Attribute function)

-   Latitude

    -   Decimal degrees

-   Longitude

    -   Decimal degrees

-   CFS

    -   Modeled median flow based on Perry et al. 2004

-   EPSG\*

    -   Datum for coordinates.

    -   \*If not provided, this function will create this column for you
        anD auto-populate it with the value 4269 (NAD83, North America)
        <https://epsg.io/4269>

If you named the required fields differently in your input file, it's
ok. Click on the drop-down menus. The list of fields in your input file
will appear. Select the correct column headings.

All columns in the input file are carried over into the output file.

# Disclaimers

The StreamCat R tools that are used to match your sites with the
NHDPlusV2 stream segment identifiers (COMIDs) do not have QC measures.
At a minimum, we recommend the following check:

-   Include waterbody name in your input file and check against
    \'gnis_name\' in the output file (which is derived from the
    NHDPlusV2 dataset); if they don\'t match, investigate.

Depending on the size of your input file, this function may several
minutes to run (e.g., 5-10 minutes). Be patient. There is a lot going on
behind the scenes. If you streamline your input file and only include
station information (vs. station + macroinvertebrate data), it will run
faster. You can then use the \'Merge Files\' function to join the
macroinvertebrate + stations data.

# Test file

Click [here](https://github.com/Blocktt/ShinyAppDocuments/tree/main/KDHEtools/Example_Files) to download an example file, R scripts, and data (all functions).

------------------------------------------------------------------------

*Last updated 2023-07-28*
