---
title: "Background Information"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
always_allow_html: true
output:
  html_fragment:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results='asis', echo=FALSE, warning=FALSE, message = FALSE)
```

## Background

This R Shiny application calculates numeric Index of Biotic Integrity (IBI) scores based on diatom samples collected by the Wyoming Department of Environmental Quality ([WDEQ](https://deq.wyoming.gov/)).

WDEQ monitors the conditions of streams and rivers in Wyoming to inform management decisions regarding protection, mitigation, and restoration of aquatic life and other designated uses. WDEQ has diatom assemblage data from nearly 700 river and stream samples collected since 2007. WDEQ's data provide an opportunity to test diatom metrics with independently established disturbance gradients and classification based on natural variability within Wyoming. The IBI consolidates measures of the aquatic community (metrics) that, when combined, indicate the similarity of a biological sample to expected conditions of a well-balanced community. The process by which the metrics are selected and combined in an index follows established and innovative analytical methods of the reference condition approach ([Hughes et al. 1986](https://link.springer.com/article/10.1007/BF01866767), [Bailey et al. 2004](https://link.springer.com/book/10.1007/978-1-4419-8885-0)). In this approach, biological conditions that are sampled from relatively undisturbed sites and that account for natural variability are set as a standard (or reference) to which other samples are compared ([Stoddard et al. 2006](https://esajournals.onlinelibrary.wiley.com/doi/10.1890/1051-0761%282006%29016%5B1267%3ASEFTEC%5D2.0.CO%3B2)). Using metrics to establish the numeric index scale, IBI values that resemble those found in reference sites are determined to meet the expectations for a well-balanced aquatic community. IBI values that are unlike the reference values indicate departure from the acceptable biological conditions and probable impairment of aquatic life uses.

The Wyoming Diatom IBI calibration project resulted in an index applicable statewide. Four metrics are adjusted by eight different factors (see Metric Adjustment Download Tab) that represent natural stream types. Reports detailing the development of the Wyoming Diatom IBI can be found [here](https://github.com/Blocktt/ShinyAppDocuments/tree/main/WDEQtools/Reports).

## Intended application

The purpose of the index is to describe the diatom community of Wyoming freshwater streams and rivers for the purpose of biological condition assessment. A diatom IBI enhances the state of Wyoming's monitoring and assessment strategy by adding a numeric indicator of diatom community structure that could be used to:

-   Evaluate attainment of Wyoming's narrative surface water quality criteria protective of aquatic life for perennial streams and rivers
-   Assist in identification of chemical and physical stressors to aquatic life of perennial streams and rivers in Wyoming
-   Identify improvements in aquatic life of perennial streams and rivers in Wyoming following restoration efforts.

The Wyoming Diatom IBI in combination with the Wyoming Stream Integrity Index (WSII) and Wyoming River InVertebrate Prediction And Classification System (RIVPACS) provides an accurate assessment of ecological effects on two aquatic assemblages (diatoms and benthic macroinvertebrates) in Wyoming. The diatom IBI might respond to different stressors compared to those indicated by the existing macroinvertebrate assemblages, and so, improves WDEQ's diagnostic ability to identify pollutants that may contribute to water quality degradation.

Sampled streams and rivers were wadeable and perennial. All diatom samples used to develop the Wyoming Diatom IBI were collected by WDEQ in accordance with the WDEQ Manual of Standard Operating Procedures ([WDEQ 2021](https://deq.wyoming.gov/water-quality/watershed-protection/water-quality-assessment/)). The Wyoming Diatom IBI should only be used to evaluate biological conditions on wadeable perennial streams and rivers in Wyoming. Moreover, diatom samples need to be collected during low flow conditions from July 15 through October. If the user applies the indices to samples that do not meet these criteria, results should be interpreted with caution. For more information on sampling methodologies and application limits, see the Wyoming Diatom IBI development reports [here](https://github.com/Blocktt/ShinyAppDocuments/tree/main/WDEQtools/Reports).

## Input metrics and scoring formulae

The Wyoming Diatom IBI was calibrated for applicability statewide; however, four of the seven metrics are adjusted by metric adjustment factors. The metric adjustment factors represent natural stream types. Table 1 shows the input metrics and scoring formulae for the IBI (metrics in need of adjustment have the "\_RFadj" suffix). See the **"Metric Adjustment Download"** tab for more information on factors and how to assign factor values to each site.

#### Table 1. Metrics and scoring formulae for the Wyoming Diatom IBI. DE = discrimination efficiency, Trend = metric response to stress, 5th = 5th percentile of metric values, and 95th = 95th percentile of metric values.

```{r table_1}
library(readxl)
library(knitr)
library(kableExtra)
# state directories
table.dir <- "tables"
table.file <- "Instruction_Tables.xlsx"
tab1.dir <- "DiatomMetrics"

table1 <- read_excel(file.path(table.dir, table.file), sheet = tab1.dir
                     , na = c("NA", ""), trim_ws = TRUE, skip = 0
                     , col_names = TRUE)  

options(knitr.kable.NA = '')
table1 %>%
  kable() %>%
  kable_styling(full_width = F, position = "left", font_size = 14)
```
