---
title: "Background on Predictor Variables"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
always_allow_html: true
output:
  html_fragment:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results='asis', echo=FALSE, warning=FALSE, message = FALSE)
```

### Metric Adjustment Factors

The Kansas Macroinvertebrate MMI contains five metrics that are adjusted based on site-specific natural factors including flow estimates from Perry et al. 2002, StreamCat estimates [(StreamCat)](https://www.epa.gov/national-aquatic-resource-surveys/streamcat-dataset), and Level III Ecoregion. To retrieve data for a site, the user must first obtain the appropriate COMID by conducting a spatial join between the site location (i.e., latitude/longitude) and the NHDPlus V2 flowline layer (outside of the app). Once each input site has been assigned a COMID, the user can use the **Prepare Data** tab to obtain COMID-specific data. Metric adjustment factor data are required for the **Calculate MMI** input file. See Table 2 for additional information on the natural variables. See Table 3 (in **Calculate MMI** tab) for necessary input fields. For more information on the derivation of these metric adjustment factors, please read the Kansas Macroinvertebrate MMI [report](https://github.com/Blocktt/ShinyAppDocuments/tree/main/KDHEtools/Reports). Note, only COMIDs from the state of Kansas can be used. If the user would like COMIDs from other, nearby states, please contact Ben Block (Ben.Block\@tetratech.com) and he can update the available list.

#### Table 2. Background information on natural variables and which macroinvertebrate metrics each adjusted.

```{r table_2}
library(readxl)
library(knitr)
library(kableExtra)
# state directories
table.dir <- "tables"
table.file <- "Instruction_Tables.xlsx"
tab2.dir <- "MetAdjFactors"

table2 <- read_excel(file.path(table.dir, table.file), sheet = tab2.dir
                     , na = c("NA", ""), trim_ws = TRUE, skip = 0
                     , col_names = TRUE)  

# kable(table2)
options(knitr.kable.NA = '')

table2 %>% 
  kbl() %>% 
  kable_styling(full_width = F, position = "left", font_size = 14)
```
