---
title: "Instructions and Background Information"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
always_allow_html: true
output:
  html_fragment:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results='asis', echo=FALSE, warning=FALSE, message = FALSE)
```

# Index Calculator Instructions

KDHEtools was developed to calculate macroinvertebrate metrics and a Multi-Metric Index (MMI) for wadeable streams in Kansas. Detailed documentation of the Kansas Macroinvertebrate MMI development can be found [here](https://github.com/Blocktt/ShinyAppDocuments/tree/main/KDHEtools/Reports). Users can run any data through the Index calculator and get a result. However, if samples do not meet the criteria listed on the **Background** page, results should be interpreted with caution.

The Instructions are divided into three sections: 1) preparing the input file, 2) Index Calculator operation, and 3) frequently asked questions (FAQ). Links to an example input file and a document containing more detailed information on preparing the input file are also provided.

## Preparing the input file

KDHEtools requires input files to be in a tabular format, saved as either comma-separated values (CSV) (most commonly used) or tab-separated (.txt or .tsv) files. Table 1 contains a list of required fields. Column names must match the spelling and symbology exactly as shown (some column headings have underscores). Note, macroinvertebrate traits are not required for each record (i.e., trait fields allow null values).

Click here to view an [EXAMPLE INPUT FILE](https://github.com/Blocktt/ShinyAppDocuments/tree/main/KDHEtools/Example_Files). If the user is unsure how to download a data file from GitHub, [watch this short video](https://www.youtube.com/watch?v=GIJdfuAoqFI) or contact Ben Block (Ben.Block\@tetratech.com).

#### Table 1. List of required fields. Column names must match the spelling and symbology exactly as shown.

```{r table_1}
library(readxl)
library(knitr)
library(kableExtra)
# state directories
table.dir <- "tables"
table.file <- "Instruction_Tables.xlsx"
tab1.dir <- "Instr_Table_Bugs"

table1 <- read_excel(file.path(table.dir, table.file), sheet = tab1.dir
                     , na = c("NA", ""), trim_ws = TRUE, skip = 0
                     , col_names = TRUE)  

# kable(table1)
table1 %>% 
  kbl() %>% 
  kable_styling(full_width = F, position = "left", font_size = 14)
```

### Metric Adjustment Factors

The Kansas Macroinvertebrate MMI contains five metrics that are adjusted based on site-specific natural factors including flow estimates from Perry et al. 2001, StreamCat estimates [(StreamCat)](https://www.epa.gov/national-aquatic-resource-surveys/streamcat-dataset), and Level III Ecoregion. To retrieve data for a site, the user must first obtain the appropriate COMID by conducting a spatial join between the site location (i.e., latitude/longitude) and the NHDPlus V2 flowline layer (outside of the app). Once each input site has been assigned a COMID, the user can use the **Metric Adjustment Download** tab to obtain COMID-specific data. Metric adjustment factor data are required for the **Index Calculator** input file. See Table 2 for additional information on the natural variables. See Table 3 for necessary input fields. For more information on the derivation of these metric adjustment factors, please read the Kansas Macroinvertebrate MMI [report](https://github.com/Blocktt/ShinyAppDocuments/tree/main/KDHEtools/Reports). Note, only COMIDs from the state of Kansas can be used. If the user would like COMIDs from other, nearby states, please contact Ben Block (Ben.Block\@tetratech.com) and he can update the available list.

#### Table 2. Background information on natural variables.

```{r table_2}
library(readxl)
library(knitr)
library(kableExtra)
# state directories
table.dir <- "tables"
table.file <- "Instruction_Tables.xlsx"
tab2.dir <- "MetAdjFactors"

table2 <- read_excel(file.path(table.dir, table.file), sheet = tab2.dir
                     , na = c("NA", ""), trim_ws = TRUE, skip = 0
                     , col_names = TRUE)  

# kable(table2)
table2 %>% 
  kbl() %>% 
  kable_styling(full_width = F, position = "left", font_size = 14)
```

#### Table 3. List of required fields. Column names must match the spelling and symbology exactly as shown.

```{r table_3}
library(readxl)
library(knitr)
library(kableExtra)
# state directories
table.dir <- "tables"
table.file <- "Instruction_Tables.xlsx"
tab3.dir <- "MetAdjFactInput"

table3 <- read_excel(file.path(table.dir, table.file), sheet = tab3.dir
                     , na = c("NA", ""), trim_ws = TRUE, skip = 0
                     , col_names = TRUE)  

# kable(table3)
table3 %>% 
  kbl() %>% 
  kable_styling(full_width = F, position = "left", font_size = 14)
```

#### Metric Adjustment Download Instructions

## Index Calculator Operation Instructions

Once the **Index Calculator** tab is opened, the user will see the Index Calculator interface. The user should follow the onscreen instructions as follows:

1.  Load file
    -   **Default settings work well**. If the user is unsure of how to format the data, simply store as a .csv and use the default load settings.
    -   An example input file can be downloaded from Github [EXAMPLE LINK](https://github.com/Blocktt/ShinyAppDocuments/tree/main/KDHEtools/Example_Files)
    -   Choose Separator. The separator indicates how the data is stored. Comma indicates the input file is a comma-separated file (.csv). Tab indicates the input file is a tab-separated file (.txt or .tsv). Finally, Semicolon indicates the input file is a semicolon-separated file which is uncommon in the United States but common in western European countries. Be certain that the designated separator is not used in any of the text fields.
    -   Choose Quote. The quote indicates how the data is stored. Double quote is the most common.
    -   Choose file to upload. Hit the browse button and search for the input file.
    -   Once uploaded, make sure data look correct in the viewer. If the incorrect Separator or Quote is chosen, the user may receive an error or the data may look incorrect in the viewer.
2.  Calculate Index
    -   Click the 'Calculate Metric Values and Scores' button.
    -   A progress bar will appear in the bottom, right corner of the screen. Results cannot be downloaded (step 3) until the progress bar is completed and disappears.
3.  Download results
    -   Select the button to download a zip file with inputs and results.
    -   Check 'results_log.txt' for any warnings or messages. Note, some warnings are automatically generated by R. Feel free to reach out to Ben Block (Ben.Block\@tetratech.com) for any questions related to warnings.

## Frequently asked questions and troubleshooting

1.  Why am I getting an error saying that I am missing columns even when I am not?
    -   You may have incorrectly spelled a given column. Also, some columns (e.g., Index_Name) require an underscore to separate the two words.
2.  Why does my data look strange in the data viewer?
    -   You likely have the incorrect Separator or Quote selected. Otherwise, there may be commas in text fields when the comma separator is selected.
3.  The Index calculation is taking forever to calculate, has the app frozen?
    -   Even though R works remarkably fast, a large dataset will cause the app to slow down. The estimate is approximately 30 seconds for small datasets; however, this could extend to a few minutes for large datasets.
