---
title: "Background Input Metrics"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
always_allow_html: true
output:
  html_fragment:
    toc: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results='asis', echo=FALSE, warning=FALSE, message = FALSE)
```

## Input metrics and scoring formulae

The Kansas Macroinvertebrate MMI was calibrated for applicability
statewide; however, all five metrics are adjusted by metric adjustment
factors. The metric adjustment factors represent natural stream types.
Table 1 shows the input metrics and scoring formulae for the MMI
(metrics in need of adjustment have the "\_RFadj" suffix). See the
**"Metric Adjustment Download"** tab for more information on factors and
how to assign factor values to each site.

#### Table 1. Metrics and scoring formulae for the Kansas Macroinvertebrate MMI. DE = discrimination efficiency, Trend = metric response to stress, 5th = 5th percentile of metric values, and 95th = 95th percentile of metric values. All metrics were adjusted by randomForest models (RF adj).

```{r table_1}
library(readxl)
library(knitr)
library(kableExtra)
# state directories
table.dir <- "tables"
table.file <- "Instruction_Tables.xlsx"
tab1.dir <- "BugMetrics"

table1 <- read_excel(file.path(table.dir, table.file), sheet = tab1.dir
                     , na = c("NA", ""), trim_ws = TRUE, skip = 0
                     , col_names = TRUE)  

options(knitr.kable.NA = '')
table1 %>%
  kable() %>%
  kable_styling(full_width = F, position = "left", font_size = 14)
```
