---
title: "Background Information"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
always_allow_html: true
output:
  html_fragment:
    toc: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results='asis', echo=FALSE, warning=FALSE, message = FALSE)
```

## Background

This R Shiny application (KDHEtools) calculates numeric Multi-Metric
Index (MMI) scores based on samples collected from wadeable streams in
Kansas (statewide). Five metrics are adjusted by 12-16 different factors
that represent natural stream types (see the **"Metric Adjustment
Download"** tab). A report detailing the development of the Kansas
Macroinvertebrate MMI can be found
[here](https://github.com/Blocktt/ShinyAppDocuments/tree/main/KDHEtools/Reports).

## Funding

Development of KDHEtools and the underlying MMI was funded by the Kansas
Department of Health and Environment (KDHE).

## Intended application

The purpose of the index is to describe the macroinvertebrate community
of Kansas freshwater streams and rivers for the purpose of biological
condition assessment. A macroinvertebrate IBI enhances the state of
Kansas's monitoring and assessment strategy by adding a numeric
indicator of macroinvertebrate community structure that could be used
to:

-   Evaluate attainment of Kansas's narrative surface water quality
    criteria protective of aquatic life for wadeable streams and rivers
-   Assist in identification of chemical and physical stressors to
    aquatic life of wadeable streams and rivers in Kansas
-   Identify improvements in aquatic life of wadeable streams and rivers
    in Kansas following restoration efforts.

The Kansas Macroinvertebrate MMI is intended to be applied to samples
that meet the following criteria:

-   Geographic area: Kansas

-   Stream type: freshwater, wadeable

-   Subsample size: 200-count samples are recommended for best
    performance; once you get below 100, interpret with caution

-   Taxonomic resolution: genus or species-level as allowed by available
    keys, specimen condition, and specimen maturity

-   Collection gear: 500-µm mesh D-frame net and fine point forceps •
    Collection period: April 15--October 15

-   Collection method: KDHE Stream Probabilistic
    ([SP](https://www.kdhe.ks.gov/1230/Stream-Probabilistic)) program or
    Stream Biological
    ([SB](https://www.kdhe.ks.gov/1233/Stream-Biological-Monitoring-Program))
    Monitoring Program protocols. Organisms are composited from multiple
    habitats. Field staff try to maximize the diversity of organisms
    being captured, targeting macrohabitats (riffles, leafpacks,
    undercut banks and rootmats, fine substrate, aquatic vegetation, and
    large woody debris). One or two-person field crews collect organisms
    with 500-µm mesh D-frame nets, using sweeps or kicks, or by hand
    picking organisms off large hard substrates. Organisms are picked in
    the field and identified in the lab, going to genus or species-level
    as allowed by available keys, specimen condition, and specimen
    maturity.

For more information on sampling methodologies and application limits,
see the Kansas Macroinvertebrate MMI development report
[here](https://github.com/Blocktt/ShinyAppDocuments/tree/main/KDHEtools/Reports).

## Input metrics and scoring formulae

The Kansas Macroinvertebrate MMI was calibrated for applicability
statewide; however, all five metrics are adjusted by metric adjustment
factors. The metric adjustment factors represent natural stream types.
Table 1 shows the input metrics and scoring formulae for the MMI
(metrics in need of adjustment have the "\_RFadj" suffix). See the
**"Metric Adjustment Download"** tab for more information on factors and
how to assign factor values to each site.

#### Table 1. Metrics and scoring formulae for the Kansas Macroinvertebrate MMI. DE = discrimination efficiency, Trend = metric response to stress, 5th = 5th percentile of metric values, and 95th = 95th percentile of metric values. All metrics were adjusted by randomForest models (RF adj).

```{r table_1}
library(readxl)
library(knitr)
library(kableExtra)
# state directories
table.dir <- "tables"
table.file <- "Instruction_Tables.xlsx"
tab1.dir <- "BugMetrics"

table1 <- read_excel(file.path(table.dir, table.file), sheet = tab1.dir
                     , na = c("NA", ""), trim_ws = TRUE, skip = 0
                     , col_names = TRUE)  

options(knitr.kable.NA = '')
table1 %>%
  kable() %>%
  kable_styling(full_width = F, position = "left", font_size = 14)
```

## Contact Information

If you encounter problems with the KDHEtools app and/or have suggestions
on ways to further improve it, please contact Ben Block from Tetra Tech
via email (Ben.Block\@tetratech.com). For more information on this
project, please contact Tony Stahl (Anthony.Stahl\@ks.gov).
